The Weakly Nonlinear Saturation of the Magnetorotational Instability

### 2014-09-12 paper/mri_paper.tex
more intro

For matter to accrete from disks onto their central objects, some mechanism for angular momentum transport beyond mere molecular viscosity must be invoked. Since its discovery by \citeauthor{Chandrasekhar:1960wh} (\citeyear{Chandrasekhar:1960wh}) and subsequent rediscovery by \citeauthor{Balbus:1991vs} (\citeyear{Balbus:1991vs}), the MRI remains the leading explanation for rapid angular momentum transport in astrophysical disks. The instability arises when a differentially rotating disk is threaded by a vertical magnetic field. The presence of the magnetic field destabilizes the disk gas, driving turbulence and angular momentum transport.

The MRI evolves via a period of linear growth, followed by nonlinear saturation. The saturation mechanism is not well understood.

@@

\bibitem[{Balbus(1991)}]{Balbus:1991vs}
Balbus, S A and Hawley, J F, The Astrophysical Journal, 376, 214

>>>>

For matter to accrete from a disk onto a central object, angular momentum must be transported radially outward in the disk. The transport mechanism is likely turbulent (\citei{Shakura:1973wg}). Since its discovery by \citeauthor{Chandrasekhar:1960wh} (\citeyear{Chandrasekhar:1960wh}) and subsequent rediscovery by \citeauthor{Balbus:1991vs} (\citeyear{Balbus:1991vs}), the MRI remains the leading explanation for rapid angular momentum transport in astrophysical disks. The instability arises when a differentially rotating disk is threaded by a vertical magnetic field. The presence of the magnetic field destabilizes the disk gas, driving turbulence and angular momentum transport.

The MRI evolves via a period of linear growth, followed by nonlinear saturation. The nonlinear evolution of the MRI is a complex process, and an understanding of the mechanism by which the MRI saturates remains elusive. Most previous works on MRI saturation have studied the problem in simulation. Indeed, only a few works have tackled MRI saturation analytically.

%more mention of other analytic works

This work builds on the weakly nonlinear analysis of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}).

@@

\bibitem[{Balbus \& Hawley(1991)}]{Balbus:1991vs}
Balbus, S A and Hawley, J F, The Astrophysical Journal, 376, 214

\bibitem[{Shakura \& Sunyaev(1973)}]{Shakura:1973wg}
Shakura, N I and Sunyaev, R A, Astronomy \& Astrophysics, 24, 337

\bibitem[{Umurhan et al.(2007)}]{Umurhan:2007hs}
Umurhan, O M, Regev, O, Menou, K, Phys. Rev. E, 76, 036310
